---
title: "Week Three"
author: "Eline Smit"
date: "5/16/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

In this case we use the median, as using the mean will cause outliers to influence the data more easily. Using the median

```{r}
library(deSolve)

setwd("/homes/kbsmit/Documents/project8/SystemsBiologyProject8/Data/")
expdata = read.csv("MPL.csv")

medians <- aggregate(expdata[,c("MPL_conc","mRNA","Free_receptor")],list(expdata$dose,expdata$time), median, na.rm=T)
names(medians)[1:2] <- c("dose","time")

#set the initial states, timeframe, and parameters, and create a function for the formulas
state <- c(rm0 = 4.74, 
           r0 = 267, 
           dr = 0, 
           drn = 0)

timeframe <- seq(0, 168, by = 1)

parmezan <- c(ksrm = 2.90, 
            ic50rm = 26.2, 
            kon = 0.00329, 
            kt = 0.63, 
            kre = 0.57, 
            Rf = 0.49, 
            kdr = 0.0572, 
            kdrm = 0.612, 
            ksr = 3.22, 
            D = 39)

formulas <- function(t, y, parameters){
  with(as.list(c(y, parameters)), {
    dmRNAr <- ksrm * (1 - (drn) / (ic50rm + drn)) - kdrm * rm0
    dR <- ksr * rm0 + Rf * kre * drn - kon * D * r0 - kdr * r0
    dDR <-kon * D * r0 - kt * dr
    dDRN <- kt * dr - kre * drn
    
    return(list(c(dmRNAr, dR, dDR, dDRN)))
    }
  )
}

# generate output from using all the data given, then adjust D for the 2nd concentration and generate another output
out  <- ode(times = timeframe, y = state,   parms = parmezan, func = formulas, method = "euler")

parmezan["D"] <- 107

out2  <- ode(times = timeframe, y = state,   parms = parmezan, func = formulas, method = "euler")

# Plot the graphs for mRNA with the first concentration first
par(yaxs = "i", xaxs = "i")
plot(out[,c("time", "rm0")], type="l", xlab = "Time in hours", ylab = "fmol/g", main = "mRNA concentration w/ dose = 0.1", xlim = c(0, 170), ylim = c(0,5))
lines(medians[medians$dose == 0.1, c("time", "mRNA")], col=2)
lines(expdata[expdata$dose == 0.1, c("time", "mRNA")], col=2, type = "p")

# Plot the graphs for mRNA with the second concentration
par(yaxs = "i", xaxs = "i")
plot(out2[,c("time", "rm0")], type="l", xlab = "Time in hours", ylab = "fmol/g", main = "mRNA concentration w/ dose = 0.3", xlim = c(0, 170), ylim = c(0,5))
lines(medians[medians$dose == 0.3, c("time", "mRNA")], col=2)
lines(expdata[expdata$dose == 0.3, c("time", "mRNA")], col=2, type = "p")

# Now do the same for the receptor concentrations
par(yaxs = "i", xaxs = "i")
plot(out[,c("time", "r0")], type="l", xlab = "Time in hours", ylab = "fmol/mg", main = "Receptor concentration w/ dose = 0.1", xlim = c(0, 170), ylim = c(0,500))
lines(medians[medians$dose == 0.1, c("time", "Free_receptor")], col=2)
lines(expdata[expdata$dose == 0.1, c("time", "Free_receptor")], col=2, type = "p")

par(yaxs = "i", xaxs = "i")
plot(out2[,c("time", "r0")], type="l", xlab = "Time in hours", ylab = "fmol/mg", main = "Receptor concentration w/ dose = 0.3", xlim = c(0, 170), ylim = c(0,500))
lines(medians[medians$dose == 0.3, c("time", "Free_receptor")], col=2)
lines(expdata[expdata$dose == 0.3, c("time", "Free_receptor")], col=2, type = "p")

```

The simulation from the model does not seem to match the experimental data very well when it comes to the mRNA concentration, while it does seem to match the model pretty well for the concentration of the free receptor, this difference between the experiment and the model could be caused by differences in parameters in the experimental data. The kd_Rm and the ks_Rm parameters seem the most likely to be affecting this difference.


# Assignment 2

[1] To remove the effect from the medicine on the rate of synthesis of mRNA, the formula $k_{s\_rm}\cdot(1-\frac{DR(N)}{IC_{50\_Rm}+DR(N)})-k_{d\_Rm}\cdot mRN\kern-0.17emA_r$ needs to be changed to only $k_{s\_rm}-k_{d\_Rm}\cdot mRN\kern-0.17emA_r$.

```{r}
parmezan["D"] <- 120*1000/374.471

edited_formulas_one <- function(t, y, parameters){
  with(as.list(c(y, parameters)), {
    dmRNAr <- ksrm - kdrm * rm0
    dR <- ksr * rm0 + Rf * kre * drn - kon * D * r0 - kdr * r0
    dDR <-kon * D * r0 - kt * dr
    dDRN <- kt * dr - kre * drn
    
    return(list(c(dmRNAr, dR, dDR, dDRN)))
    }
  )
}
out  <- ode(times = timeframe, y = state,   parms = parmezan, func = formulas, method = "euler")
plot(out, lty = 1, main = c("mRNA concentration", "Receptor concentration", "Receptor complex concentration", "Receptor complex conc. in nucleus"), xlab = "Time in hours", ylab = c("fmol/g", "fmol/mg", "fmol/mg", "fmol/mg"))

```


\newpage
# Appendix: All code for this report

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
